---
output: github_document
---

```{r setup, include = FALSE}
auth_success <- tryCatch(
  cancerscreening:::khis_cred_docs(),
  cancerscreening_cred_internal_error = function(e) e
)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE,
  purl = cancerscreening::khis_has_cred(),
  eval = cancerscreening::khis_has_cred()
)
```

# cancerscreening

## Overview

The goal of `cancerscreening` is to provide a easy way to download cancer screening data from the Kenya Health Information System (KHIS) using R.

## Installation

You can install the released version of cancerscreening from [CRAN](https://cran.r-project.org/) with:
```r
install.packages("cancerscreening")
```

You can install the development version of from [Github](https://github.com) with:
```r
# install.packages("devtools")
devtools::install_github('damurka/cancerscreening')
```
___Note: This package is not yet available on CRAN.___

## Usage

### Load cancerscreening package

```{r}
library("cancerscreening")
```

### Package conventions

* Most function begin with the prefix `get_` followed by the screening area `cervical` or `breast`. Auto-completion is your friend
* Goal is to allow the download of data associated with the data of interest, e.g. `get_cervical_screened`, `get_cervical_positive`, or `get_cervical_treated`
* cancerscreening is "pipe-friendly" and, infact rexports `%>%` but does not require its use.

### Quick demo

Before calling any function that makes an API call you need credentials to [KHIS](https://hiskenya.org). You will be expected to set this credential to download the data. The credentials are stored securely with the operating system using the package `keyring`

```{r eval = FALSE}
# Set the credentials
khis_cred(username = 'KHIS username')
```

After setting the credential you can invoke any function to download data from the API

```{r}
# Download the cervical cancer screening data for country
cacx_screened <- get_cervical_screened('2022-07-01')
head(cacx_screened)
```

### Metadata reuse

When you need data from more than one function it is efficient to download the
metadata and share among the functions as shown below:

```{r}
# Download the organisation units
organisations <- get_organisation_units_metadata()
head(organisations)

# Download category options
categories <- get_category_options_metadata()
head(categories)

# Download data elements 
elements <- get_data_elements_metadata()
head(elements)

# Download cervical cancer screening data
cacx_screened <- get_cervical_screened('2021-07-01',
                                       end_date = '2021-12-31',
                                       level = 'county',
                                       elements = elements,
                                       categories = categories,
                                       organisations = organisations)
head(cacx_screened)

# Download cervical cancer screening positives
cacx_positive <- get_cervical_positive('2021-07-01',
                                       end_date = '2021-12-31',
                                       level = 'county',
                                       elements = elements,
                                       categories = categories,
                                       organisations = organisations)
head(cacx_positive)

# Download Breas mammogram screening
breast_mammogram <- get_breast_mammogram('2021-07-01', 
                                         end_date = '2021-12-31',
                                         level = 'county',
                                         elements = elements,
                                         categories = categories,
                                         organisations = organisations)
head(breast_mammogram)
```

