---
title: "Screening Coverage"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE
)
```

Screening overage refers to the extent of participation of eligible (i.e., target age) women in the screening program in a given time period and is calculated by dividing the number of eligible women screened during a given time by the total number of eligible women. High coverage of the target population is one of the most important components of a successful cancer prevention program (Pretorius et al. 1991, Sasieni 1991, WHO 1992).

### Cervical Cancer Screening Coverage

The number of women screened can be obtained from `get_cervical_screened` and the eligible women can be obtained from the `get_cervical_target_population`.

```{r eval=FALSE}
# Target population by county
target_population <- get_cervical_target_population(2023, level = 'county')

# Population screened during the same period
screened <- get_cervical_screened('2023-01-01', level = 'county')

# aggregate the screened 
screened <- screened %>%
  summarise(screened = sum(value, na.rm = TRUE), .by = county)

# merge with the target and calculate the coverage

screened %>%
  right_join(target_population, by = 'county') %>%
  mutate(coverage = screened / target)
```

### CBE Coverage

The number of women screened can be obtained from `get_breast_cbe` and the eligible women can be obtained from the `get_breast_cbe_target_population`.

```{r eval=FALSE}
# Target population by county
target_population <- get_breast_cbe_target_population(2023, level = 'county')

# Population screened during the same period
screened <- get_breast_cbe('2023-01-01', level = 'county')

# aggregate the screened 
screened <- screened %>%
  summarise(screened = sum(value, na.rm = TRUE), .by = county)

# merge with the target and calculate the coverage

screened %>%
  right_join(target_population, by = 'county') %>%
  mutate(coverage = screened / target)
```
